Experiment 0 pilot

### 2018-07-20 experiments/e0/writeup/e0.Rmd
e0 is a mess

```{r"make a tibble with factors explicit"}
# make a df to carry forward
d <- tibble(
  obs = din$obs, 
  trial = din$trial, 
  subject = as.factor(din$subject), 
  item = factor(din$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9")),
  gap = factor(din$gap, levels = c("BigGap", "SmallGap"), labels = c("Big Gap", "Small Gap")), 
  quan = factor(din$target, levels = c("BigTarget", "SmallTarget"), labels = c("Big Target", "Small Target")), 
  vag = factor(din$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")), 
  subit = factor(ifelse(din$pair %in% c(1,2,5,6), "Subitizable", "Not subitizable")),
  side = factor(din$side), 
  text = factor(din$Text),
  err = 1 - din$isCorrect, 
  rt_raw = ifelse(din$rt >=0, din$rt, NA),
  rt_ms = ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA)), # restricting rt to positive values (rt>=0) converted a score of -6 ms to NA
  rt_log = log( ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA)) ),
  "rt_reciprocal[-1/rt]" = -1/ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA)), # We multiplied reciprocal scores [1/RT] by -1 [-1/RT] to maintain the direction of effects compatible for the three variants, effectively converting speed [number of responses per ms] into 'rate of slowing' - A linear mixed model analysis of masked repetition priming, Kliegl, Masson & Richter, VISUAL COGNITION,2010,18(5),655--681 (p.~662)
  rt_reciprocal = -1000/ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA))
)
d$c.vag = ifelse(d$vag=="Crisp", -.5, .5)
```

Size of the data is `r nrow(d)` trials. 1 was removed for being impossible ($-6~ms$). There are `r nrow(subset(d,err==1))` error trials. The variable rt_ms has NA in the case that an error was made. The variable rt_raw includes rt for error trials.

@@

### Uncapped, max of 20,000 ms

@@

data_for_plot <- gather(d, key=rt_metric, value=rt_value, rt_ms, rt_log, "rt_reciprocal[-1/rt]") %>%
  arrange(rt_value)

ga1untrimmed = ggplot(data_for_plot, aes(x=rep(1:1600,times=3), y=rt_value)) + 
  geom_point(na.rm=TRUE) + 
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free_y") +
  xlab("observations, ordered by rt_value") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

ga2untrimmed = ggplot(data_for_plot, aes(rt_value)) +
  geom_density(na.rm=TRUE) +
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

@@

### Capped at 10,000 ms

```{r fig.show='hide', warning=FALSE, echo=FALSE}
# now with a simple cap imposed at 10000 ms, half-way through the maximum allotted time of 20000 (2 minute timeout)
summary(d$rt_ms[d$rt_ms<10000])

data_for_plot <- gather(d %>% filter(rt_ms<10000), key=rt_metric, value=rt_value, rt_ms, rt_log, "rt_reciprocal[-1/rt]") %>%
  arrange(rt_value)

ga1trimmed = ggplot(data_for_plot, aes(x=rep(1:1556,times=3), y=rt_value)) + 
  geom_point(na.rm=TRUE) + 
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free_y") +
  xlab("observations, ordered by rt_value")+ theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

ga2trimmed = ggplot(data_for_plot, aes(rt_value)) +
  geom_density(na.rm=TRUE) +
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r selectTrim}
dat <- d %>% filter(rt_ms<10000)
```

@@

```{r, lmers}
lmer2 <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item))
# lmer3 is closer to the original anova than lmer2
lmer3 <- tidy(lme4::lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item)))
dwplot(lmer3 %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2)) + xlab("Coefficient")
```

>>>>

```{r "make a tibble with factors explicit"}
# make a df to carry forward
d <- tibble(
  obs = din$obs, 
  trial = din$trial, 
  subject = as.factor(din$subject), 
  item = factor(din$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9")),
  gap = factor(din$gap, levels = c("BigGap", "SmallGap"), labels = c("Big Gap", "Small Gap")), 
  quan = factor(din$target, levels = c("BigTarget", "SmallTarget"), labels = c("Big Target", "Small Target")), 
  vag = factor(din$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")), 
  subit = factor(ifelse(din$pair %in% c(1,2,5,6), "Subitizable", "Not subitizable")),
  side = factor(din$side), 
  text = factor(din$Text),
  err = 1 - din$isCorrect, 
  rt_ms = din$rt,
  rt_log = log( din$rt ),
  rt_reciprocal = -1/din$rt # rate of slowing: # We multiplied reciprocal scores [1/RT] by -1 [-1/RT] to maintain the direction of effects compatible for the three variants, effectively converting speed [number of responses per ms] into 'rate of slowing' - A linear mixed model analysis of masked repetition priming, Kliegl, Masson & Richter, VISUAL COGNITION,2010,18(5),655--681 (p.~662) 
)
```

```{r "make centred variables for factors"}
d$c.gap   <- ifelse(d$gap=="Big Gap", -.5, .5)
d$c.quan  <- ifelse(d$quan=="Big Target", -.5, .5)
d$c.vag   <- ifelse(d$vag=="Crisp", -.5, .5)
d$c.subit <- ifelse(d$subit=="Subitizable", -.5, .5)
d$c.side  <- ifelse(d$side=="Left", -.5, .5)
```

Size of the data is `r nrow(d)` trials.  There are `r nrow(subset(d,err==1))` error trials. Error trials are removed from analysis of RT

```{r}
d <- subset(d, err==0)
```

<!-- 1 was set to NA for being impossible ($-6~ms$). -->
<!-- The variable rt_ms has NA in the case that an error was made. The variable rt_raw includes rt for error trials. -->

@@

### Uncapped

@@

data_for_plot <- d %>% 
  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
  arrange(rt_value)

ga1untrimmed = ggplot(data_for_plot, aes(x=rep(1:(nrow(data_for_plot)/3),times=3), y=rt_value)) + 
  geom_point(na.rm=TRUE) + 
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free_y") +
  xlab("observations, ordered by rt_value") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

ga2untrimmed = ggplot(data_for_plot, aes(rt_value)) +
  geom_density(na.rm=TRUE) +
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

@@

### Capped at 0 ms and 10,000 ms

```{r fig.show='hide', warning=FALSE, echo=FALSE}
# now with a simple cap imposed to select RTs between 0 and 10000 ms
summary(d$rt_ms[d$rt_ms >= 0 & d$rt_ms<10000])

data_for_plot <- d %>% 
  filter(rt_ms >= 0 & rt_ms <= 10000) %>% 
  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
  arrange(rt_value)

ga1trimmed = ggplot(data_for_plot, aes(x=rep(1:(nrow(data_for_plot)/3), times=3), y=rt_value)) + 
  geom_point(na.rm=TRUE) + 
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free_y") +
  xlab("observations, ordered by rt_value")+ theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

ga2trimmed = ggplot(data_for_plot, aes(rt_value)) +
  geom_density(na.rm=TRUE) +
  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r selectTrim}
dat <- d %>% filter(rt_ms >= 0 & rt_ms <= 10000) 
```

```{r plot rt per trial by subject}
dat %>% 
  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
  filter(rt_metric=="rt_log") %>%
  ggplot(aes(x=trial, y=rt_value, col=vag)) +
  geom_point() +
  geom_smooth(se=F) +
  facet_wrap(~subject) +
  theme(aspect.ratio = 1, panel.background = element_rect(fill="white", colour="black", linetype = "solid"))
```

```{r}
dat %>% 
  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
  filter(rt_metric=="rt_log") %>%
  ggplot(aes(group=vag, y=rt_value)) +
  geom_boxplot(aes(x=item, col=vag))
```

@@

```{r, lmer1}
lmer1.ms  <- lmer(data=dat, rt_ms ~         vag + subit + quan + gap + side + (1|subject) + (1|item))
lmer1.log <- lmer(data=dat, rt_log ~        vag + subit + quan + gap + side + (1|subject) + (1|item))
lmer1.rcp <- lmer(data=dat, rt_reciprocal ~ vag + subit + quan + gap + side + (1|subject) + (1|item))

models <- 
rbind(
  cbind(tidy(lmer1.ms),model="ms"), 
  cbind(tidy(lmer1.log),model="log"), 
  cbind(tidy(lmer1.rcp),model="rcp")
)

dwplot(models, vline = geom_vline(xintercept = 0)) + facet_wrap(~model, scales="free", ncol=1)

```

```{r mler on logrt}
lmer2.log <- lmer(data=dat, rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item))
dwplot(lmer2.log %>% tidy() %>% filter(group=="fixed")) + 
  ggtitle("rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)")

lmer2 <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item))
# lmer3 is closer to the original anova than lmer2
lmer3 <- tidy(lme4::lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item)))
dwplot(lmer3 %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2)) + xlab("Coefficient")
```
